Finite-state morphological transducers for three Kypchak languages

### 2013-10-14 kipchak-lrec-2013/kipchak-paper.tex
more notes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48037 72bbbca6-d526-0410-a7d9-f06f51895060
more work
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48079 72bbbca6-d526-0410-a7d9-f06f51895060

\name{Jonathan North Washington, Ilnar Salimzyanov, Francis M. Tyers}

\address{Departments of Linguistics and Central Eurasian Studies\\
Indiana University\\
Bloomington, IN 47405 (USA)\\
\texttt{jonwashi@indiana.edu} \and
Institut für Maschinelle Sprachverarbeitung \\
Universität Stuttgart\\
Stuttgart (Germany) \\
\texttt{ilnar@ilnar} \and 
Departament de Llenguatges i Sistemes Informàtics \\  
Universitat d'Alacant\\
E-03071 Alacant (Spain)\\
\texttt{ftyers@dlsi.ua.es} 
}

@@

\cite{bammatov1960}

@@

@@

\begin{table}
\begin{center}
\begin{tabular}{lrrr}
		\hline
\multirow{2}{*}{\textbf{Part of speech}} & \multicolumn{3}{|c|}{\textbf{Number of stems}} \\
                        & Kazakh & Tatar & Kumyk \\
		\hline
		Noun & - & - & - \\
		Verb & - & - & - \\
		Adjective & - & - & - \\
		Proper noun & - & - & - \\
		Adverb & - & - & - \\
		Numeral & - & - & - \\
		Conjunction & - & - & - \\
		Postposition & - & - & - \\
		Pronoun & - & - & - \\
		Determiner & - & - & - \\
		\hline
		Total: & - & - & - \\
		\hline
\end{tabular}
\end{center}

@@

\begin{table}
\begin{center}
\begin{tabular}{lrr}
\textbf{Corpus} & \textbf{Words} & \textbf{Coverage} \\
\hline
wikipedia & 850K &  - \\
Äwezov & 155K & - \\
RFERL 2010 & 3.2M & - \\
bible & 577K & - \\
quran & 107K & - \\
\hline
\end{tabular}
 \caption{Corpora used for coverage of Kazakh}
 \label{table:corpora-kaz}
\end{center}
\end{table}

@@

%Code switching
%More languages

@@

>>>>

\name{Jonathan North Washington$^\dagger$, Ilnar Salimzyanov$^\ddagger$, Francis M. Tyers$^\star$}

\address{$^\dagger$Departments of Linguistics and Central Eurasian Studies\\
Indiana University\\
Bloomington, IN 47405 (USA)\\
\texttt{jonwashi@indiana.edu} \and
$^\ddagger$Institut für Maschinelle Sprachverarbeitung \\
Universität Stuttgart\\
Stuttgart (Germany) \\
\texttt{ilnar@ilnar} \and 
$^\star$Departament de Llenguatges i Sistemes Informàtics \\  
Universitat d'Alacant\\
E-03071 Alacant (Spain)\\
\texttt{ftyers@dlsi.ua.es} 
}

@@

\cite{bammatov1960}
\cite{olmesov2000}

@@

% Noun morphotactics -- is basically equivalent in all three, except for archiphonemes
% Adjective categorisation
% Adverb categorisation
% "Irregular" harmony, caused by orthography
% Non-finite verb form categorisation "converbs"

% Handling numerals and acronyms.

\subsection{Development effort}
% how long it took
% make a graph for the kumyk one

\subsection{Statistics}

\begin{table}
\begin{center}
\begin{tabular}{|l|rrr|}
		\hline
\multirow{2}{*}{\textbf{Part of speech}} & \multicolumn{3}{|c|}{\textbf{Number of stems}} \\ \cline{2-4}
                        & Kazakh & Tatar & Kumyk \\
		\hline
		Noun & - & - & - \\
		Verb & - & - & - \\
		Adjective & - & - & - \\
		Proper noun & - & - & - \\
		Adverb & - & - & - \\
		Numeral & - & - & - \\
		Conjunction & - & - & - \\
		Postposition & - & - & - \\
		Pronoun & - & - & - \\
		Determiner & - & - & - \\
		\hline
		Total: & - & - & - \\
		\hline
\end{tabular}
 \caption{Number of stems in each of the categories}
 \label{table:coverage}
\end{center}

@@

We have evaluated the morphological analysers in two ways. The first was by calculating the naïve coverage and mean ambiguity 
on freely available corpora. The second was by performing an evaluation of precision and recall on some 
smaller, hand-validated test sets.

\subsection{Corpora}

@@

@@

%Code switching - +example
%More languages: Nogai, Bashkir, Karakalpak, Karachay-Balkar

@@

\section*{Acknowledgements}

We would like to thank the Google Code-in (2011) for supporting the development 
of the Kazakh transducer and the Google Summer of Code (2012) for supporting the 
development of both the Kazakh and the Tatar transducers.
